Cycles of the logistic map

### 2013-05-03 doc/pc-ijbc.tex
small edits

\newcommand{\vx}{\vct x}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

A polynomial is cyclic if it is invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a(\vx) = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b(\vx) = x_1 x_3 + x_2 x_4$,
  where $\vx \equiv \{x_1, \ldots, x_n\}$ and $n = 4$.
Note that a cyclic polynomial is not necessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a(\vx)$ and $b(\vx)$ are not symmetric,
  but $a(\vx) + b(\vx)$ is.

A cyclic polynomial can be generated by summing over
  distinct cyclic versions of a simpler polynomial of $x_k$, or a \emph{generator}.
For example,
$x_1 x_2$ is a generator of $a(\vx)$;
$x_1 x_3$ is that of $b(\vx)$;
and
$x_1 x_2 + \frac{1}{2} x_1 x_3$ is that of $a(\vx) + b(\vx)$.
Note that the coefficient before $x_1 x_3$
  is 1 in the second case
  for there are only two distinct versions,
  but is $\frac{1}{2}$ in the third case
  for there are four. % distinct versions of the generator.

Cyclic polynomials
  generated from monomials of unit coefficient
  [such as $a(\vx)$ and $b(\vx)$, but not $a(\vx) + b(\vx)$]
  can be labeled as follows.
We first write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  then the sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s
  serves as the label of the cyclic polynomial $C_p(\vx)$.
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
The cycle length $n$ is omitted for convenience.
%  for we will mostly work with a fixed $n$ at a time.
In case the cyclic polynomial has several generators,
%  e.g., both $x_1 x_2$ and $x_2 x_3$ are generators of $C_{12}(\vx)$
%   (assuming $n \ge 3$),
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vct x = \{x_1, \ldots, x_n\}$
  is a linear combination of
the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}

\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since the original polynomial is cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
%Since no square or higher powers of $x_k$
%  can survive the reduction,
%  the final cyclic polynomials are square-free;
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only variable introduced by the substitutions.
\end{proof}

For example, for $n = 2$, the cyclic polynomial
  $K(\vx) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K(\vx) = (R + 1) \, C_{1}(\vx) - 2 R \, C_0(\vx)$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx)$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients then must have a zero determinant,
  which yields the desired polynomial.

@@

For $n = 1$, we have two square-free cyclic polynomials
\big[omitting ``$(\vx)$''\big]
  $C_0 = 1$ and $C_1 = x_1$;
  and $\Lambda = -2 \, x_1$
Thus,
$\Lambda \, C_0 = -2 \, x_1$,
$\Lambda \, C_1 = -2 \, x_1^2 = -2 \, R + 2  \, x_1$,
and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$, the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
0=  \left| \begin{array}{cc}
  \lambda     & 2          \\
  2 R         & \lambda - 2
  \end{array} \right| = -4R - 2\lambda + \lambda^2.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
Thus, $R = (\lambda^2 -2\lambda)/4$;
  and the fixed point begins at $R_a = -1/4$ ($\lambda = +1$),
  and becomes unstable at $R_b = 3/4$ ($\lambda = -1$).

@@

If an $n$-cycle is not born out of the above branching,
  we call it an \emph{original} cycle,
The two types of cycles exist
  as separate factors of $P_n(R, \lambda = 1)$.
One can then show that the factor $S_n(R)$ responsible for the original cycles
  %or the \emph{original factor} below,
  is given by
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      \left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      \right]
    },
  \label{eq:origfac}
  \end{equation}
where the inner product on the denominator is carried over
   $k$ from 1 to $c$ that are coprime to $c$.

Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$,
  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
  the onset of the original cycle.
Note $R_a' = 5/4$ is excluded from $S_4(R)$
  as it comes from period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).

>>>>

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx^{(#1)}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

A polynomial is cyclic if it is invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ (for $n = 4$).
%
Note that a cyclic polynomial is not necessarily symmetric
  (i.e., invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  but $a + b$ is.

A cyclic polynomial can be generated by summing over
  distinct cyclic versions of a simpler polynomial of $x_k$, or a \emph{generator}.
For example,
$x_1 x_2$ is a generator of $a$;
$x_1 x_3$ is that of $b$;
and
$x_1 x_2 + \frac{1}{2} x_1 x_3$ is that of $a + b$.
Note that the coefficient before $x_1 x_3$
  is 1 in the second case
  for there are only two distinct versions,
  but $\frac{1}{2}$ in the third case
  for there are four. % distinct versions of the generator.

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We first write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  then the sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s
  serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$.
We will, however, drop the superscript ``$^{(n)}$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear.
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%The cycle length $n$ is omitted for convenience.
%  for we will mostly work with a fixed $n$ at a time.
In case the cyclic polynomial has several generators,
%  e.g., both $x_1 x_2$ and $x_2 x_3$ are generators of $C_{12}(\vx)$
%   (assuming $n \ge 3$),
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx)$
  formed by the $n$-cycle points
  $\vx = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vx)$:
\[
  K(\vx) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx)$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since the original polynomial is cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx) \equiv 1$.
%Since no square or higher powers of $x_k$
%  can survive the reduction,
%  the final cyclic polynomials are square-free;
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vxn{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients then must have a zero determinant,
  which yields the desired polynomial.

@@

For $n = 1$, we have two square-free cyclic polynomials
  $C_0 = 1$ and $C_1 = x_1$;
  and $\Lambda = -2 \, x_1$
Thus,
$\Lambda \, C_0 = -2 \, x_1$,
$\Lambda \, C_1 = -2 \, x_1^2 = -2 \, R + 2  \, x_1$,
and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$, the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
0=  \left| \begin{array}{cc}
  \lambda     & 2          \\
  2 R         & \lambda - 2
  \end{array} \right| = -4R - 2\lambda + \lambda^2.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
Thus, $R = (\lambda^2 -2\lambda)/4$;
  and the fixed point begins at $R_a = -1/4$ ($\lambda = +1$),
  and becomes unstable at $R_b = 3/4$ ($\lambda = -1$).

@@

If an $n$-cycle is not born out of the above branching,
  we call it \emph{original}.
The two types of cycles exist
  as separate factors of $P_n(R, \lambda = 1)$.
One can then show that the factor $S_n(R)$ responsible for the original cycles
  %or the \emph{original factor} below,
  is given by
  \begin{equation}
    S_n(R)
    = \frac
    {
      P_n(R, 1)
    }
    {
      \prod_{c d =  n, \; c > 1}
      %\left[ \,
        \prod_{(k, c) = 1}
      P_{d}
        \left(
          R, e^{2k\pi i/c}
        \right)
      %\right]
    },
  \label{eq:origfac}
  \end{equation}
where the inner product on the denominator is carried over
   $k$ from 1 to $c$ that are coprime to $c$.

Take the $n = 4$ case as an example, we have
$\prod_{k=1,3} P_1(R, e^{k\pi i/2})
%=(-4R-2i-1)(-4R+2i-1)
=16R^2+8R+5$
(for $c = 4$),
$P_2(R, -1) = 4R - 5$
(for $c = 2$),
and
$P_4(R, 1)
= (4R - 5) (16 R^2 + 8 R + 5)
  \bigl[
    (4R - 3)^3 - 108
  \bigr]$.
Dividing $P_4(R, 1)$ by the two factors yields
$S_4(R) = (4R-3)^3 - 108$,
  whose only real root $R_a=(3+\sqrt[3]{108})/4$ corresponds to
  the onset of the original cycle.
Note $R_a' = 5/4$ is excluded from $S_4(R)$
  as it is due to period-doubling the 2-cycle.
The two cycles are compared in \reffigs{cobweb}(d) and (e).
